---
title: "STATS 201/8 Assignment 5"
author: "Kabir Patel Kpat913"
date: 'Due Date: 3pm Thursday 12th October'
output:
  html_document:
    fig_caption: yes
    number_sections: yes
  pdf_document: 
    number_sections: yes
  word_document: 
    number_sections: yes
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.height=3)
```

```{r setup, echo=FALSE, warning=FALSE}
## Do not delete this!
## It loads the s20x library for you. If you delete it 
## your document may not compile
library(s20x)
```

# Question 1

## Questions of Interest

It was of interest to compare e-mail habits of adults, depending on their marital status.

## Read in and inspect the data:

```{r,fig.height=4.6,fig.width=6}
email.df = read.csv("email.csv",stringsAsFactors = TRUE)
summary(email.df)
with(email.df, table(email))  
plot(email~marital,data=email.df)
```

## Comment on plot(s) and/or summary statistics

We can see here that the marital status of divorced had a count that was significantly less than the other two groups and that the never group had a count nearly twice the divorced group had. The married group also had a count more than twice what divorced had.

## Fit and check model

```{r}
Emailmarital.gfit = glm(email~marital, family = poisson, data=email.df)
plot(Emailmarital.gfit, which=1)
summary(Emailmarital.gfit)
1 - pchisq(35637, 3966)

Emailmarital.gfit2 = glm(email~marital, family = quasipoisson, data=email.df)
plot(Emailmarital.gfit2, which=4)
summary(Emailmarital.gfit2)

exp(confint(Emailmarital.gfit2))

email.df$marital = relevel(email.df$marital, ref="married")
emailmarital.gfit3 = glm(email~marital, family = quasipoisson, data = email.df)
exp(confint(emailmarital.gfit3))

email.df$marital = relevel(email.df$marital, ref="never")
emailmarital.gfit4 = glm(email~marital, family = quasipoisson, data = email.df)
exp(confint(emailmarital.gfit4))

anova(Emailmarital.gfit2, test="F")






```

## Method and Assumption Checks

As the response variable, emails, is a count, we have fitted a generalized linear model with a quasipoisson response distribution. We have one explanatory variable with three factors, Married, Divorced, Never.

All assumptions were satisfied, there was evidence on the poisson model as the p-value was 0 which is lower then 0.05 so we can conclude that model was not adequate as there was overdispersion, thus using the quasipoisson model.

For our final model, we assume that the email count for observation is poisson with mean $u_i$ where

$log(u_i) = \beta_0 + \beta_1 \times maritalmarried_i + \beta_2 \times maritalnever_i$

and where $maritalmarried_i$ and $maritalnever_i$ are dummy variables which take the value 1 if observation is from $maritalnever_i$, otherwise they are 0.

We see that the marital status is significant as the Anova shows there is a difference between marital status because the p-value is lower than 0.05.

## Executive Summary

We wanted to know if there was a difference in the number of emails sent by the 3 different marital groups, (married, divorced, never).

There is evidence of a difference in emails sent between the three groups, but the most significant difference is between married and divorced.

We estimate that on average the divorced group send between 2.48% and 36.93% higher than someone with in the married group.

------------------------------------------------------------------------

# Question 2

## Questions of Interest

It was of interest to see how increasing APACHE II scores related to the mortality rate. In particular, to assess the effect of a 1 point or 5 point difference in APACHE II scores and to estimate mortality rates for APACHE II scores of 10, 20, 30 and 40.

### Read in and inspect the data:

```{r,fig.height=4.6,fig.width=6}
Apache.df=read.csv("Apache.csv",header=TRUE)
plot(deaths/number_of_patients~apache_score, 
     ylab="prob. of death", 
     xlab="APACHE II score",data=Apache.df)
head(Apache.df)
tail(Apache.df,10)

```

## Comment on plot

We see that as the Apache II score increases the prob. of death increases as well. Notable S curve.

\## Fit and check model

```{r}
Apache.fit=glm(deaths/number_of_patients~apache_score, family = binomial, data= Apache.df, weights = number_of_patients)
plot(Apache.fit, which = 1)
plot(Apache.fit, which = 4)
summary(Apache.fit)
1-pchisq(43.999, 36)
exp(confint(Apache.fit))
exp(5*confint(Apache.fit))

```

## Get intervals to predict the probability of a patient admitted to ICU dying with 30 days with APACHE II scores of 10, 20, 30 or 40.

```{r}
predn=predictGLM(Apache.fit, data.frame(apache_score=c(10,20,30,40)), type="response")
predn

     
```

## Method and Assumption Checks

The data recorded the number of patients and deaths where we then got the probability of death from deaths/number of patients. We then fitted a Binomial GLM with a single predictor of Alpache score (numeric).

There was no evidence of over dispersion as (p-Value = 0.169) so we can trust the results from this binomial model.

Our final model:

$log(prob_i) = \beta_0 + \beta_1 \times Alpache_i$

where prob (probability of death) is deaths/number of patients.

## Write sentences, as if for an Executive Summary, interpret your four predictions intervals.

The probability of a patient admitted to ICU dying with 30 days with APACHE II score of 10 is between 0.196 and 0.298. This increases to between 0.446 and 0.565 for a score of 20, between 0.668 and 0.840 for a score of 30 and between 0.826 and 0.957 for a score of 40.

## Write sentences, as if for an Executive Summary, quantifying the effect of a 1 point and 5 point increase in APPACHE II scores.

We estimate that for every 1 point increase in the APPACHE II scores, the probability of death increases by between 1.089 - 1.159 and with a 5 point increase the probability of death increases by between 1.530 and 2.095.

## If we overlaid the model on the original plot, what shape would it be? (See question sheet for options.)

S-Shaped Curve
